---
title       : Data mining Final presentation
subtitle    : Analysis of German Credit Data
author      : Abdul Sadhiq K M
job         : Business analyst
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : zenburn      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
github:
  user: abdulsadhiq
  repo: Datamining
---
<section class="slide1">
  <h2>German Credit Data - Project description</h2>
  <p>When a bank receives a loan application, based on the applicant's profile the bank has to make a decision regarding whether to go ahead with the loan approval or not</p>
  <ol>
    <li>
      <h3>Risk factors</h3>
      <p>If the applicant is a good credit risk, i.e. is likely to repay the loan, then not approving the loan to the person results in a loss of business to the bank</p>
	  <p>If the applicant is a bad credit risk, i.e. is not likely to repay the loan, then approving the loan to the person results in a financial loss to the bank</p>
    </li>
    <li>
      <h3>Overall objective</h3>
      <p>Minimization of risk and maximization of profit on behalf of the bank</p>
    </li>	
  </ol>
</section>

---
<h2>Environment set-up</h2>
```{r simple-plot, echo=T, fig.align='center', fig.height=6, message=T}
options(warn = -1)
library(klaR)
library(MASS)
data("GermanCredit")
library(caret)
require(ggplot2)
GermanCredit.org<-GermanCredit
```

--- 
```{r echo = F, echo=TRUE, fig.width=12, message=TRUE, warning=FALSE, results='asis'}
#levels(GermanCredit.org$personal_status_sex)
barchart(table(GermanCredit.org$personal_status_sex))
```

--- 
<h2>Removing the additional level from Gender</h2>
```{r echo = F, echo=TRUE, fig.width=12, message=TRUE, warning=FALSE, results='asis'}
GermanCredit.org$personal_status_sex<-droplevels(GermanCredit.org$personal_status_sex)
levels(GermanCredit.org$personal_status_sex)<-levels(GermanCredit.org$personal_status_sex)[1:4]
levels(GermanCredit.org$personal_status_sex)
```

---
```{r echo = F, echo=TRUE, fig.width=12, message=TRUE, warning=FALSE, results='asis'}
library(plotrix)
pie3D(table(GermanCredit.org$personal_status_sex),
    labels = levels(GermanCredit.org$personal_status_sex), 
main = "An exploded 3D pie chart", explode=0.1, radius=.9, labelcex = 1.3,  start=0.7)
```


---
```{r echo = F, echo=TRUE, fig.width=12, message=TRUE, warning=FALSE, results='asis'}
ggplot(GermanCredit.org, aes(age)) + geom_histogram(binwidth=4, colour="black", fill="green") +
  labs(x= "Age",y= "Frequency" , title = "Plot of Age")
```

---
```{r echo = F, echo=TRUE, fig.width=12, message=TRUE, warning=FALSE, results='asis'}

ggplot(GermanCredit.org,aes(x=age,fill=credit_risk))+facet_wrap(~personal_status_sex)+
  geom_histogram(binwidth = 5)+labs(x="Customer Age",y="Customer count",
     title="Age distribution by gender and marital status")
```

---
```{r echo = F, echo=TRUE, fig.width=12, message=TRUE, warning=FALSE, results='asis'}

ggplot(GermanCredit.org,aes(x=age,fill=credit_risk))+
  facet_wrap(~housing)+
  geom_histogram(binwidth = 5)+labs(x="Customer Age",y="Customer count",
          title="Age distribution by housing")
```

---
```{r echo = F, echo=TRUE, fig.width=12, message=F, warning=FALSE, results='asis'}
ggplot(GermanCredit.org,aes(x=age,fill=credit_risk))+
  facet_wrap(housing~personal_status_sex)+
    geom_histogram(binwidth = 5)+labs(x="Customer Age",y="Customer count",
title="Age distribution by gender , marital status and housing")


```

---
<h2>Actual data is biased towards Good</h2>
```{r echo = F, echo=TRUE, fig.width=12, message=TRUE, warning=FALSE, results='asis'}
qplot(GermanCredit.org$credit_risk,main="Raw data proportion")
```

---
<h2>...hence balancing the output variable</h2>
```{r echo = T, fig.width=12, message=F, warning=FALSE, results='asis'}
#install.packages("unbalanced")
library(unbalanced)
library(mice)
#prop.table(table(GermanCredit.org$credit_risk))
n<-ncol(GermanCredit.org)
output <-GermanCredit.org[,n]
input <- GermanCredit.org[,-n]
data1<-ubBalance(X=input,Y=output,type="ubSMOTE",positive="bad",percOver=35, 
                 percUnder=450)
overdata1<-data.frame(data1$X, credit_risk=data1$Y)
#Removing null values that were introduced after balancing
#md.pattern(overdata1)
overdata1<-overdata1[complete.cases(overdata1),]
#prop.table(table(overdata1$credit_risk))
```

---
```{r echo = F, echo=FALSE, fig.width=12, message=F, warning=FALSE, results='asis'}
par(mfrow=c(1,2))
pie3D(table(GermanCredit.org$credit_risk),
      labels =round(prop.table(table(GermanCredit.org$credit_risk)),2),
      main = "Before under sampling", explode=0.1, radius=.9, labelcex = 1.3,
      start=0.7)

pie3D(table(overdata1$credit_risk),
      labels = round(prop.table(table(overdata1$credit_risk)),2),
      main = "After under sampling", explode=0.1, radius=.9, labelcex = 1.3,
      start=0.7)
par(xpd=TRUE)
legend(1,0.7,legend=levels(overdata1$credit_risk),yjust=0.2, xjust = -0.1,
       fill = heat.colors(overdata1$credit_risk))
```

---
<h2>Random partition for train and test data set after balancing</h2>

```{r, echo=TRUE, message=TRUE, warning=FALSE}

indx1<-(sample(1:nrow(overdata1),floor(nrow(overdata1)/3)))
train_data<-overdata1[-indx1,]
test_data<-overdata1[indx1,]
nrow(train_data)#67%
nrow(test_data)#33%
nrow(overdata1[rownames(test_data)%in%rownames(train_data),]) #Validation
```

---
<h2>Butilding decision tree to identify the important features of the the data set that are actually contributing to our output variable</h2>
```{r, echo=TRUE, message=FALSE, warning=FALSE}
par(mfrow=c(1,1))
rp1<-rpart::rpart(credit_risk~.,data=train_data)
rpart::printcp(rp1) # Returns important variables
```

---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot(rp1)
text(rp1, use.n=TRUE, all=TRUE, cex=.8)
```

---
<h1>Apply the identified important featrures to each models</h1>
<h2>Model 1 : GLM</h2>
```{r, echo=TRUE, message=TRUE, warning=FALSE}
model_glm_f<-caret::train(credit_risk~age+amount+credit_history+duration+
                            installment_rate+housing+
                            property+purpose+status,
                          data=train_data,method="glm",family="binomial")
predict_glm_f<-predict(model_glm_f,newdata = test_data)
conf_glm<-confusionMatrix(data=predict_glm_f,test_data$credit_risk)
p1<-conf_glm$overall['Accuracy']
p1
```

---
<h2>Model 2 : Random Forest</h2>

```{r, echo=TRUE, message=FALSE, warning=FALSE}
model_nb_f<-caret::train(credit_risk~age+amount+credit_history+duration+
                           installment_rate+housing+
                           property+purpose+status,data=train_data,method="rf")
predict_nb_f<-predict(model_nb_f,newdata = test_data)
conf_rf<-confusionMatrix(data = predict_nb_f,test_data$credit_risk)
p2<-conf_rf$overall['Accuracy']
p2
```

---
<h2>Model 3 : Naive Bayes</h2>
```{r, echo=TRUE, message=TRUE, warning=FALSE}
model_nb_f<-caret::train(credit_risk~age+amount+credit_history+duration+
                             installment_rate+housing+
                             property+purpose+status,data=train_data,method="nb")
predict_nb_f<-predict(model_nb_f,newdata = test_data)
conf_nb<-confusionMatrix(data = predict_nb_f,test_data$credit_risk)
p3<-conf_nb$overall["Accuracy"]
p3
```

---
<h2>Accuray comparision to identify the winner!!!</h2>
```{r, echo=TRUE, message=TRUE, warning=FALSE}
Final_comp<-c(glm=p1,rf=p2,nb=p3)
dotplot(Final_comp,horizontal=T)
```